---
output: github_document
always_allow_html: true
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE,
  fig.retina = 2,
  fig.align = 'center'
)
```

# watercostaccra

<!-- badges: start -->

[![License: CC BY
4.0](https://img.shields.io/badge/License-CC_BY_4.0-pink.svg)](https://creativecommons.org/licenses/by/4.0/)

<!-- badges: end -->

The goal of watercostaccra is to ...

## Installation

You can install the development version of watercostaccra from
[GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("openwashdata/watercostaccra")
```

Alternatively, you can download the individual data sets as a CSV or XLSX
file from the table below.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(knitr)
library(readr)
library(stringr)
library(gt)
library(kableExtra)
library(ggtext)

extdata_path <- "https://github.com/openwashdata/watercostaccra/raw/main/inst/extdata/"

read_csv("data-raw/dictionary.csv") |> 
  distinct(file_name) |> 
  dplyr::mutate(file_name = str_remove(file_name, ".rda")) |> 
  dplyr::rename(dataset = file_name) |> 
  mutate(
    CSV = paste0("[Download CSV](", extdata_path, dataset, ".csv)"),
    XLSX = paste0("[Download XLSX](", extdata_path, dataset, ".xlsx)")
  ) |> 
  knitr::kable()

```

## Data

The package provides access to ...

```{r}
library(watercostaccra)
```

The `watercostaccra1` data set contains data about a household survey on water costs and coping strategies in Accra. It has `r nrow(watercostaccra1)` observations and `r ncol(watercostaccra1)` variables. The `watercostaccra2` data set contains data about a water point survey conducted in Accra as well. It has `r nrow(watercostaccra2)` observations and `r ncol(watercostaccra2)` variables. For an overview of the variable names, see the following table.

```{r echo=FALSE, message=FALSE, warning=FALSE}
readr::read_csv("data-raw/dictionary.csv") |>
  dplyr::filter(file_name == "watercostaccra1.rda") |>
  dplyr::select(variable_name:description) |> 
  knitr::kable() |> 
  kableExtra::kable_styling("striped")
```

## Example

Here is an example illustrating 

```{r}
library(watercostaccra)


```


## License

Data are available as
[CC-BY](https://github.com/openwashdata/watercostaccra/blob/main/LICENSE.md).

## Citation

Please cite this package using:

```{r}
citation("watercostaccra")
```
